---
title: "S01-generate_sim_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{S01-generate_sim_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lextremagam)
library(tidyr)
library(dplyr)
library(ggplot2)
library(arrow)
```

# Treatment variables 
Generating the treatment variables.

Predictor (x) range : 0 to 1, by increments of 0.05 (21 unique predictor values)
Testing the effect of sample size (n): 21, 42, 63, 84, 105

## Experiment 1: testing the effect of sample size (sampling effort) on peak detection capacity
5 different sample size treatment levels ( n = 21, 42, 63, 84, 105), representing 1, 2, 3, 4, and 5 replicate measures respectively.

```{r }
x <- seq(0,1,length = 21)
rep <- 1000

# Experiment 1 ####
func <- c("norm", "linear")
param <-0.5
n <- c(21, 42, 63, 84, 105)
dup <- c(1, 2, 3, 4, 5)
error <- 0.75 #DOUBLE CHECK
list_names <- c("n_21", "n_42", "n_63", "n_84", "n_105")

sim_dat_e1 <- expand_grid(func = func,
                          param = param,
                          rep = 1:rep,
                          error = error,
                          dup = rep(dup, times= dup),
                          x = x)

set.seed(1)
sim_dat_e1 <- sim_dat_e1 <- sim_dat_e1%>%
  dplyr::group_by(func, param, error, rep, dup)%>%
  dplyr::mutate(sample_id = cur_group_id(), 
                scenario_id = paste0(func,"_p", param, "_n", dup, "_e", error))%>%
  dplyr::mutate(y_true_unscaled = case_when(
    func=="norm"~dnorm(x, mean = param, sd = 4/20),
    func=="linear"~x))%>%
  dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
  dplyr::mutate(y_true = y_true/max(y_true))%>%
  dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
  dplyr::ungroup()

#Export data to parquet
setwd("~/Documents/Pedersen2023/Peaks/lextremagam")
write_parquet(x = sim_dat_e1, "data-raw/sim/data/sim_dat_e1.parquet")

sim_dat_e1_plot <- ggplot(data=filter(sim_dat_e1, rep==1))+
  geom_point(aes(x=x, y=y, colour=error), size=0.1)+
  geom_smooth(aes(x=x, y=y, colour=error))+
  facet_wrap(~dup)
sim_dat_e1_plot


```


## Experiment 2: testing the effect of residual error level (signal strength) on peak detection capacity
19 different error levels, and 1000 replicate samples per error level treatment. Based on a normally distributed curve centered in the range with a standard deviation of .2.

```{r }
x <- seq(0,1,length = 21)
rep <- 1000

# Experiment 2 ####
func <- c("norm", "linear")
param <- 0.5
n <- 63
dup <- 2
error <- seq(0.05, 0.95, by=0.05)

sim_dat_e2 <- expand_grid(func = func,
                          param = param,
                          rep = 1:rep,
                          error = error,
                          dup = rep(dup, times= dup),
                          x = x)

set.seed(2)
sim_dat_e2 <- sim_dat_e2 <- sim_dat_e2%>%
  group_by(rep)%>%
  dplyr::group_by(func, param, error, rep)%>%
  dplyr::mutate(sample_id = cur_group_id(), 
                scenario_id = paste0(func,"_p", param, "_n", dup, "_e", error))%>%
  dplyr::mutate(y_true_unscaled = case_when(
    func=="norm"~ dnorm(x, mean = param, sd = 4/20),
    func=="linear"~ x))%>%
  dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
  dplyr::mutate(y_true = y_true/max(y_true))%>%
  dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
  dplyr::ungroup()


sim_dat_e2_plot <- ggplot(data=filter(sim_dat_e2, rep==1))+
  geom_point(aes(x=x, y=y, colour=error), size=0.2)+
  geom_line(aes(x=x, y=y_true))
sim_dat_e2_plot
```
## Experiment 3: Effect of curve shape and sample size on peak detection capacity
5 different sample sizes, 3 different curve shapes (peak width, peak location and skew), and 1000 replicate samples per sample curve shape level treatment.

```{r }
x <- seq(0,1,length = 21)
n <- c(21, 42, 63, 84, 105)
dup <- c(1, 2, 3, 4, 5)
rep <- 1000
list_names <- c("n_21", "n_42", "n_63", "n_84", "n_105")

# Experiment 3 ####
error <- 0.5 ######DOULBE CHECK
param <- seq(0.05, .95, by= .05)

func <- c("var-p-location", "var-p-skew", "var-p-width")

sim_dat_e3 <- expand_grid(func = func,
                          param = param,
                          dup = rep(dup, times= dup),
                          error = error,
                          rep = 1:rep,
                          x = x )

set.seed(3)
sim_dat_e3 <- sim_dat_e3%>%
  dplyr::group_by(func, param, error, rep, dup)%>%
  dplyr::mutate(sample_id = cur_group_id(),
                scenario_id = paste0(func,"_p", param, "_n", dup, "_e", error))%>%
  dplyr::mutate(y_true_unscaled = case_when(
    func=="var-p-location"~10*dnorm(x, mean = param, sd = 4/20),
    func=="var-p-skew"~dbeta(x, shape1=10*param, shape2=10*(1-param)),
    func=="var-p-width"~(param/(2*1*(1/(param*10)))*exp(-((abs(pred-0.5)/1)^(param*10)))))) %>%
  #10*dnorm(x,mean=.5,sd = param^2)))%>% #####DOUBLE CHEKC THIS SCALING
  dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
  dplyr::mutate(y_true = y_true/max(y_true))%>%
  dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
  dplyr::ungroup()

ggplot(data=filter(sim_dat_e3, rep==1, dup==1))+
  #geom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))+
  facet_grid(dup~func)

ggplot(data=filter(sim_dat_e3, rep==1, func=="var-p-width"))+
  #geom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))
  #facet_wrap(~as.factor(param))


# 
# 
# b <- param*10-0.4
# 
# test_gen <- expand_grid(a = a, 
#                         mu = mu,
#                         b = b,
#                         pred = pred)%>%
#   mutate(resp=b/(2*a*(1/b))*exp(-((abs(pred-mu)/a)^b)))%>%
#   group_by(a, mu, b)%>%
#   mutate(resp = resp-min(resp))%>%
#   mutate(resp = resp/max(resp))%>%
#     ungroup()
# 
# ggplot(data=test_gen)+
#   geom_line(aes(x = pred, y= resp, colour = as.factor(b)))
# 
# b <- 2^(param*2)
# a <- 1
# mu <- .5
# pred <- seq(0,1,length = 21)
# for (b in b){
#   resp <- b/(2*a*(1/b))*exp(-((abs(pred-mu)/a)^b)) 
#   resp <- resp-min(resp)
#   resp <- resp/(max(resp))
#   plot(resp~pred)
# }
```

##Experiment 4: The effect of curve shape but for non-peaked curves 

```{r }
x <- seq(0,1,length = 21)
n <- c(21, 42, 63, 84, 105)
dup <- c(1, 2, 3, 4, 5)
rep <- 1

# Experiment 4 ####
error <- 0.7 ######TO BE CHANGED
param <- seq(0.05, .95, by= .05)

func <- c("var-location", "var-scale", "var-curvature")

sim_dat_e4 <- expand_grid(func = func,
                          param = param,
                          rep = 1:rep,
                          dup = rep(dup, times=dup),
                          error = error,
                          x = x)

set.seed(4)
sim_dat_e4 <- sim_dat_e4%>%
  dplyr::group_by(func, param, error, rep)%>%
  dplyr::mutate(sample_id = cur_group_id(), 
                scenario_id = paste0(func,"_p", param, "_n", dup, "_e", error))%>%
  dplyr::mutate(y_true_unscaled = dplyr::case_when(
    func=="var-location"~plogis(x, location=param, scale = .1),
    func=="var-scale"~plogis(x, location=0.5, scale = (param)^3),
    func=="var-curvature"~plogis(x, location=.05, scale = (param)^3)))%>% 
  dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
  dplyr::mutate(y_true = y_true/max(y_true))%>%
  dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
  dplyr::ungroup()%>%
  dplyr::filter(func=="var-scale" & param %in% c(.6, .65, .7, .75, .8, .85, .9, .95))

ggplot(data=filter(sim_dat_e4, rep==1))+
  #ggeom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))+
  facet_wrap(~func)

ggplot(data=filter(sim_dat_e4, rep==1, func=="var-location"))+
 # geom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))+
  facet_wrap(~as.factor(param))
```

```{r }

# reps <- 1000
#
# param <- seq(0.05, .95, by= .05)
#
# func <- c("var-location", "var-scale", "var-curvature",
#           "var-p-location", "var-skew", "var-p-width")
#
# sim_data <-  expand_grid(func = func,
#                                 param=param,
#                                 x=x)
# 
#
#
# sim_data <- dplyr::group_by(sim_data,
#                             func, param)
# sim_data <- dplyr::mutate(sim_data,
#                           y_true_scaled = y_true_unscaled-min(y_true_unscaled))
# sim_data <- dplyr::mutate(sim_data,
#                           y_true_scaled = y_true_scaled/max(y_true_scaled))
# sim_data <- dplyr::ungroup(sim_data,
#                             func, param)
```
