---
title: "S01-generate_sim_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{S01-generate_sim_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lextremagam)
library(tidyr)
library(dplyr)
library(ggplot2)
```

# Treatment variables 
Generating the treatment variables.

Predictor (x) range : 0 to 1, by increments of 0.05 (21 unique predictor values)
Testing the effect of sample size (n): 21, 42, 63, 84, 105

## Experiment 1: testing the effect of sample size (sampling effort) on peak detection capacity
5 different sample size treatment levels ( n = 21, 42, 63, 84, 105), representing 1, 2, 3, 4, and 5 replicate measures respectively.

```{r }
x <- seq(0,1,length = 21)
rep <- 1000

# Experiment 1 ####
func <- "norm"
param <-0.5
n <- c(21, 42, 63, 84, 105)
dup <- c(1, 2, 3, 4, 5)
error <- 0.75 #DOUBLE CHECK
list_names <- c("n_21", "n_42", "n_63", "n_84", "n_105")

sim_dat_e1 <- expand_grid(func = func,
                          param = param,
                          rep = 1:rep,
                          error = error,
                          dup = rep(dup, times= dup),
                          x = x)

set.seed(1)
sim_dat_e1 <- sim_dat_e1 <- sim_dat_e1%>%
  dplyr::group_by(func, param, error, rep)%>%
  dplyr::mutate(sample_id = cur_group_id(), 
                scenario_id = paste0(func,"_p", param, "_n", dup))%>%
  dplyr::mutate(y_true_unscaled = dnorm(x, mean = param, sd = 4/20))%>%
  dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
  dplyr::mutate(y_true = y_true/max(y_true))%>%
  dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
  dplyr::ungroup()

sim_dat_e1_plot <- ggplot(data=filter(sim_dat_e1, rep==1))+
  geom_point(aes(x=x, y=y, colour=error), size=0.1)+
  geom_line(aes(x=x, y=y_true))+
  geom_smooth(aes(x=x, y=y, colour=error))+
  facet_wrap(~dup)
sim_dat_e1_plot
```


## Experiment 2: testing the effect of residual error level (signal strength) on peak detection capacity
19 different error levels, and 1000 replicate samples per error level treatment. Based on a normally distributed curve centered in the range with a standard deviation of .2.

```{r }
x <- seq(0,1,length = 21)
rep <- 1000

# Experiment 2 ####
func <- "norm"
param <- 0.5
n <- 63
dup <- 2
error <- seq(0.05, 0.95, by=0.05)

sim_dat_e2 <- expand_grid(func = func,
                          param = param,
                          rep = 1:rep,
                          error = error,
                          dup = rep(dup, times= dup),
                          x = x)

# sim_dat_e1 <- list()
# for (i in dup){
#   sim_dat_e1[[i]] <- expand_grid(func = func,
#                                  param = param,
#                                  rep = 1:rep,
#                                  error = error,
#                                  dup = 1:dup[i],
#                                  x = x)
# }
# names(sim_dat_e1) <- list_names

set.seed(2)
sim_dat_e2 <- sim_dat_e2 <- sim_dat_e2%>%
    group_by(rep)%>%
    dplyr::group_by(func, param, error, rep)%>%
    dplyr::mutate(sample_id = cur_group_id(), 
                  scenario_id = paste0(func,"_p", param, "_e", error))%>%
    dplyr::mutate(y_true_unscaled = dnorm(x, mean = param, sd = 4/20))%>%
    dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
    dplyr::mutate(y_true = y_true/max(y_true))%>%
    dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
    dplyr::ungroup()


sim_dat_e2_plot <- ggplot(data=filter(sim_dat_e2, rep==1))+
  geom_point(aes(x=x, y=y, colour=error), size=0.2)+
  geom_line(aes(x=x, y=y_true))
sim_dat_e2_plot
```
## Experiment 3: Effect of curve shape and sample size on peak detection capacity
5 different sample sizes, 3 different curve shapes (peak width, peak location and skew), and 1000 replicate samples per sample curve shape level treatment.

```{r }
x <- seq(0,1,length = 21)
n <- c(21, 42, 63, 84, 105)
dup <- c(1, 2, 3, 4, 5)
rep <- 1000
list_names <- c("n_21", "n_42", "n_63", "n_84", "n_105")

# Experiment 3 ####
error <- 0.5 ######DOULBE CHECK
param <- seq(0.05, .95, by= .05)

func <- c("var-p-location", "var-p-skew", "var-p-width")

sim_dat_e3 <- expand_grid(func = func,
                          param = param,
                          dup = rep(dup, times= dup),
                          error = error,
                          rep = 1:rep,
                          x = x )

# sim_dat_e2 <- list()
# for (i in dup){
#   sim_dat_e2[[i]] <- expand_grid(func = func,
#                                  param = param,
#                                  dup = 1:dup[i],
#                                  error = error,
#                                  rep = 1:rep,
#                                  x = x)
# }
# names(sim_dat_e2) <- list_names 

set.seed(3)
sim_dat_e3 <- sim_dat_e3%>%
    dplyr::group_by(func, param, error, rep, dup)%>%
    dplyr::mutate(sample_id = cur_group_id(),
                  scenario_id = paste0(func,"_p", param, "_n", dup))%>%
    dplyr::mutate(y_true_unscaled = case_when(
      func=="var-p-location"~10*dnorm(x, mean = param, sd = 4/20),
      func=="var-p-skew"~dbeta(x, shape1=10*param, shape2=10*(1-param)),
      func=="var-p-width"~10*dnorm(x,mean=.5,sd = param^2)))%>% #####DOUBLE CHEKC THIS SCALING
    dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
    dplyr::mutate(y_true = y_true/max(y_true))%>%
    dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
    dplyr::ungroup()



  
#   lapply(sim_dat_e2, function(x) {
#   x <- x%>%
#     dplyr::group_by(func, param, error, rep)%>%
#     dplyr::mutate(sample_id = cur_group_id(), 
#                   scenario_id = paste0(func,"_p", param, "_e", error))%>%
#     dplyr::mutate(y_true_unscaled = case_when(func=="var-p-location"~10*dnorm(x,mean = param, sd = 4/20),
#                                               func=="var-p-skew"~dbeta(x, shape1=10*param, shape2=10*(1-param)),
#                                               func=="var-p-width"~10*dnorm(x,mean=.5,sd = param^2)))%>% #####DOUBLE CHEKC THIS SCALING
#     dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
#     dplyr::mutate(y_true = y_true/max(y_true))%>%
#     dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
#     dplyr::ungroup()
# 
# })

ggplot(data=filter(sim_dat_e3, rep==1, dup==1))+
  #geom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))+
  facet_grid(dup~func)

ggplot(data=filter(sim_dat_e3, rep==1, func=="var-p-width"))+
  geom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))+
  facet_wrap(~as.factor(param))



```

##Experiment 4: The effect on 

```{r }
x <- seq(0,1,length = 21)
n <- c(21, 42, 63, 84, 105)
dup <- c(1, 2, 3, 4, 5)
rep <- 1000

# Experiment 3 ####
error <- 0.3 ######TO BE CHANGED
param <- seq(0.05, .95, by= .05)

func <- c("var-location", "var-scale", "var-curvature", "saturating")

sim_dat_e3 <- list(n_21 = expand_grid(func = func,
                                             param = param,
                                             rep = 1:rep,
                                             dup = 1:dup[1],
                                             error = error,
                                             x = x),
                   n_42 = expand_grid(func = func,
                                             param = param,
                                             rep = 1:rep,
                                             dup = 1:dup[2],
                                             error = error,
                                             x = x),
                   n_63 = expand_grid(func = func,
                                             param = param,
                                             rep = 1:rep,
                                             dup = 1:dup[3],
                                             error = error,
                                             x = x),
                   n_84 = expand_grid(func = func,
                                             param = param,
                                             rep = 1:rep,
                                             dup = 1:dup[4],
                                             error = error,
                                             x = x),
                   n_105 = expand_grid(func = func,
                                             param = param,
                                             rep = 1:rep,
                                             dup = 1:dup[5],
                                             error = error,
                                             x = x))


set.seed(3)
sim_dat_e3 <- lapply(sim_dat_e3, function(x) {
  x <- x%>%
    dplyr::group_by(func, param, error, rep)%>%
    dplyr::mutate(sample_id = cur_group_id(), 
                  scenario_id = paste0(func,"_p", param, "_e", error))%>%
    dplyr::mutate(y_true_unscaled = dplyr::case_when(func=="var-location"~plogis(x, location=param, scale = .1),
                                                      func=="var-scale"~plogis(x, location=0.5, scale = (param*0.7)^2),
                                                      func=="var-curvature"~plogis(x, location=.05, scale = (param*0.7)^2),
                                                     func=="saturating"~11*param*x/(1+(11*param-1)*2*x)))%>% #####DOUBLE CHEKC THIS SCALING
    dplyr::mutate(y_true = y_true_unscaled-min(y_true_unscaled))%>%
    dplyr::mutate(y_true = y_true/max(y_true))%>%
    dplyr::mutate(y = y_true + rnorm(n = n(),mean = 0, sd = error))%>%
    dplyr::ungroup()

})

ggplot(data=filter(sim_dat_e3$n_21, rep==1))+
  #ggeom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))+
  facet_wrap(~func)

ggplot(data=filter(sim_dat_e3$n_21, rep==1, func=="var-location"))+
 # geom_point(aes(x=x, y=y, colour=as.factor(param)))+
  geom_line(aes(x=x, y=y_true, colour=as.factor(param)))+
  facet_wrap(~as.factor(param))


# reps <- 1000
#
# param <- seq(0.05, .95, by= .05)
#
# func <- c("var-location", "var-scale", "var-curvature",
#           "var-p-location", "var-skew", "var-p-width")
#
# sim_data <-  expand_grid(func = func,
#                                 param=param,
#                                 x=x)
# 
#
#
# sim_data <- dplyr::group_by(sim_data,
#                             func, param)
# sim_data <- dplyr::mutate(sim_data,
#                           y_true_scaled = y_true_unscaled-min(y_true_unscaled))
# sim_data <- dplyr::mutate(sim_data,
#                           y_true_scaled = y_true_scaled/max(y_true_scaled))
# sim_data <- dplyr::ungroup(sim_data,
#                             func, param)
```
