---
title: "S04-analyze_real_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{S04-analyze_real_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lextremagam)
library(mgcv)
library(gratia)
library(dplyr)
knitr::opts_knit$set(root.dir = "~/Documents/Pedersen2023/Peaks/lextremagam")
```

# Halpern and Spies ####
```{r}
HalpernSpies1995_w1 <- read.csv("data-raw/real_data/CLEAN_HalpernSpies1995_Fig1d_w1.csv")
plot(HalpernSpies1995_w1$heterogeneity~HalpernSpies1995_w1$t_disturbance)

HalpernSpies1995_w2 <- read.csv("data-raw/real_data/CLEAN_HalpernSpies1995_Fig1d_w2.csv")
plot(HalpernSpies1995_w2$heterogeneity~HalpernSpies1995_w2$t_disturbance)

#quadratic test
hs1995w1_quad <- lextremagam:::peaktest_quad(x = HalpernSpies1995_w1$t_disturbance, 
                                             y = HalpernSpies1995_w1$heterogeneity)
#Two-lines test
hs1995w1_tl <- lextremagam:::peaktest_twolines(x = HalpernSpies1995_w1$t_disturbance, 
                                             y = HalpernSpies1995_w1$heterogeneity)
#lextremagam
hs1195w1_gam <- gam(heterogeneity~s(t_disturbance, bs="ad", k=10, m=3), method="REML", data=HalpernSpies1995_w1)
hs1195w1_lextr <- lextremagam::quantify_lextrema(mod = hs1195w1_gam, 
                                                 var = "t_disturbance", 
                                                 step_size = 0.02, 
                                                 conf_level = 0.95, 
                                                 deriv_method = "gratia", frequentist = TRUE)

hs1195w1_lextr_plot <- plot_lextrema(quant_segments = hs1195w1_lextr, plot_deriv = TRUE)
hs1195w1_lextr$segment_summary


```

#Wilson and Keddy ####
```{r}

WilsonKeddy1988 <- read.csv("data-raw/real_data/CLEAN_WilsonKeddy1988.csv")
WilsonKeddy1988$ln_somc_perc <- log(WilsonKeddy1988$somc_perc)
plot(WilsonKeddy1988$n_sp~WilsonKeddy1988$ln_somc_perc)

#quadratic test
wk1988_quad <- lextremagam:::peaktest_quad(x = WilsonKeddy1988$ln_somc_perc, 
                                             y = WilsonKeddy1988$n_sp)
#Two-lines test
wk1988_tl <- lextremagam:::peaktest_twolines(x = WilsonKeddy1988$ln_somc_perc, 
                                             y = WilsonKeddy1988$n_sp)
#lextremagam
wk1988_gam <- gam(n_sp~s(ln_somc_perc, bs="ad", k=20, m=3), method="REML", data=WilsonKeddy1988, family=nb)
wk1988_lextr <- lextremagam::quantify_lextrema(mod = wk1988_gam, 
                                                 var = "ln_somc_perc", 
                                                 conf_level = 0.95, 
                                                step_size = 0.0002,
                                                 deriv_method = "gratia")

wk1988_lextr_plot <- plot_lextrema(quant_segments = wk1988_lextr, plot_deriv = TRUE)
wk1988_lextr$segment_summary
```


#Bishop2021 Diatoms ####
```{r}
bishop2021 <- read.csv("data-raw/real_data/NBP1701_diatom_growth_rate_data_Award1543245_20220627.csv")
plot(bishop2021$specific_growth_rate~bishop2021$temperature)

bishop2021
#quadratic test
b2021_quad <- lextremagam:::peaktest_quad(x = bishop2021$temperature, 
                                             y = bishop2021$specific_growth_rate)
#Two-lines test
b2021_tl <- lextremagam:::peaktest_twolines(x = bishop2021$temperature, 
                                             y = bishop2021$specific_growth_rate)
#lextremagam
   #I think makes more sense to look at each species separately
bishop2021$species <- as.factor(bishop2021$species)
b2021_gam <- gam(specific_growth_rate~s(temperature, k=8)+s(lat, k=3), method="REML", data=dplyr::filter(bishop2021, species == "Actinocyclus curvatulus"))

b2021_lextr_ac <- lextremagam::quantify_lextrema_multivar(mod = b2021_gam, 
                                               var = "temperature", 
                                               step_size=0.015,
                                               conf_level = 0.95, 
                                               deriv_method = "gratia", 
                                               smooth = "s(temperature)")

b2021_lextr_plot <- plot_lextrema(quant_segments = b2021_lextr_ac, plot_deriv = TRUE)


```

#Sasaki et al. 
Data from "Seasonally variable thermal performance curves prevent adverse effects of heatwaves" by Matthew Sasaki, Michael Finiguerra, Hans G. Dam
11 November 2024
https://doi-org.lib-ezproxy.concordia.ca/10.1111/1365-2656.14221
https://github.com/ZoopEcoEvo/LIS_heatwave/blob/main/Output/Data/comb_d.csv
```{r}

Sasaki2024_comb_d <- read.csv("data-raw/real_data/Sasaki2024_comb_d.csv")

#one model per species per metric
# tonsa ####
  ## EPR ####
    # mod_epr_tons <- mgcv::gam(rate ~ s(temp, k = 5)+s(curve_id, bs="re"), 
    #                     data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "EPR", species == "tonsa"), curve_id = as.factor(curve_id)), method="REML")

  #this is the best one   
    mod_epr_tons <- mgcv::gam(rate ~ s(temp, k = 5, by=curve_id), 
                        data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "EPR", species == "tonsa"), curve_id = as.factor(curve_id)), method="REML")
    
  #this would be good, but lextremagam not set up for it yet
   # mod_epr_tons <- mgcv::gam(rate ~ s(temp,curve_id, bs="fs", k = 5), 
   #                        data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "EPR", species == "tonsa"), curve_id = as.factor(curve_id)), method="REML")

  lextr_epr_tons_all <- list()
  for(i in 1:length(unique(mod_epr_tons$model$curve_id))){
      lextr_epr_tons_i <- lextremagam::quantify_lextrema_multivar(mod = mod_epr_tons,
                                                                  var = "temp", 
                                                                  smooth = paste0("s(temp):curve_id", unique(mod_epr_tons$model$curve_id)[i]), 
                                                                  conf_level = 0.95, deriv_method = "gratia", frequentist = FALSE )
    lextr_epr_tons_all[[i]] <- lextr_epr_tons_i
    names(lextr_epr_tons_all)[i] <- as.character(unique(mod_epr_tons$model$curve_id)[i])
  }
  
  #extracting peaks 
  epr_tons_max <- purrr::map(lextr_epr_tons_all, ~ dplyr::filter(.x$segment_summary, feature %in% c("boundary_max", "local_max")))
  epr_tons_max <- purrr::map(epr_tons_max, ~ dplyr::mutate(.x, midpoint = (temp_end+temp_start)/2))

  
  ## HF ####
    #this is the best one   
      mod_hf_tons <- mgcv::gam(rate ~ s(temp, k = 5, by=curve_id), 
                          data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "HF", species == "tonsa"), curve_id = as.factor(curve_id)), method="REML")

    lextr_hf_tons_all <- list()
    for(i in 1:length(unique(mod_hf_tons$model$curve_id))){
        lextr_hf_tons_i <- lextremagam::quantify_lextrema_multivar(mod = mod_hf_tons,
                                                                    var = "temp", 
                                                                    smooth = paste0("s(temp):curve_id", unique(mod_epr_tons$model$curve_id)[i]), 
                                                                    conf_level = 0.95, deriv_method = "gratia", frequentist = FALSE )
      lextr_hf_tons_all[[i]] <- lextr_hf_tons_i
      names(lextr_hf_tons_all)[i] <- as.character(unique(mod_hf_tons$model$curve_id)[i])
    }
    
    #extracting peaks 
    hf_tons_max <- purrr::map(lextr_hf_tons_all, ~ dplyr::filter(.x$segment_summary, feature %in% c("boundary_max", "local_max")))
    hf_tons_max <- purrr::map(hf_tons_max, ~ dplyr::mutate(.x, midpoint = (temp_end+temp_start)/2))
    
##RF ####
 #this is the best one   
      mod_rf_tons <- mgcv::gam(rate ~ s(temp, k = 5, by=curve_id), 
                          data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "RF", species == "tonsa"), curve_id = as.factor(curve_id)), method="REML")

    lextr_rf_tons_all <- list()
    for(i in 1:length(unique(mod_rf_tons$model$curve_id))){
        lextr_hf_tons_i <- lextremagam::quantify_lextrema_multivar(mod = mod_rf_tons,
                                                                    var = "temp", 
                                                                    smooth = paste0("s(temp):curve_id", unique(mod_epr_tons$model$curve_id)[i]), 
                                                                    conf_level = 0.95, deriv_method = "gratia", frequentist = TRUE )
      lextr_rf_tons_all[[i]] <- lextr_hf_tons_i
      names(lextr_rf_tons_all)[i] <- as.character(unique(mod_hf_tons$model$curve_id)[i])
    }
    
    #extracting peaks 
    rf_tons_max <- purrr::map(lextr_rf_tons_all, ~ dplyr::filter(.x$segment_summary, feature %in% c("boundary_max", "local_max")))
    rf_tons_max <- purrr::map(rf_tons_max, ~ dplyr::mutate(.x, midpoint = (temp_end+temp_start)/2))

    
# Hudsonica ####
  ## EPR ####
  #this is the best one   
    mod_epr_huds <- mgcv::gam(rate ~ s(temp, k = 5, by=curve_id), 
                        data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "EPR", species == "hudsonica"), curve_id = as.factor(curve_id)), method="REML")

  lextr_epr_huds_all <- list()
  for(i in 1:length(unique(mod_epr_huds$model$curve_id))){
      lextr_epr_huds_i <- lextremagam::quantify_lextrema_multivar(mod = mod_epr_huds,
                                                                  var = "temp", 
                                                                  smooth = paste0("s(temp):curve_id", unique(mod_epr_huds$model$curve_id)[i]), 
                                                                  conf_level = 0.95, deriv_method = "gratia", frequentist = FALSE )
    lextr_epr_huds_all[[i]] <- lextr_epr_huds_i
    names(lextr_epr_huds_all)[i] <- as.character(unique(mod_epr_huds$model$curve_id)[i])
  }
  
  #extracting peaks 
  epr_huds_max <- purrr::map(lextr_epr_huds_all, ~ dplyr::filter(.x$segment_summary, feature %in% c("boundary_max", "local_max")))
  epr_huds_max <- purrr::map(epr_huds_max, ~ dplyr::mutate(.x, midpoint = (temp_end+temp_start)/2))

  
  ## HF ####
    #this is the best one   
      mod_hf_huds <- mgcv::gam(rate ~ s(temp, k = 5, by=curve_id), 
                          data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "HF", species == "hudsonica"), curve_id = as.factor(curve_id)), method="REML")

    lextr_hf_huds_all <- list()
    for(i in 1:length(unique(mod_hf_huds$model$curve_id))){
        lextr_hf_huds_i <- lextremagam::quantify_lextrema_multivar(mod = mod_hf_huds,
                                                                    var = "temp", 
                                                                    smooth = paste0("s(temp):curve_id", unique(mod_epr_huds$model$curve_id)[i]), 
                                                                    conf_level = 0.95, deriv_method = "gratia", frequentist = FALSE )
      lextr_hf_huds_all[[i]] <- lextr_hf_huds_i
      names(lextr_hf_huds_all)[i] <- as.character(unique(mod_hf_huds$model$curve_id)[i])
    }
    
    #extracting peaks 
    hf_huds_max <- purrr::map(lextr_hf_huds_all, ~ dplyr::filter(.x$segment_summary, feature %in% c("boundary_max", "local_max")))
    hf_huds_max <- purrr::map(hf_huds_max, ~ dplyr::mutate(.x, midpoint = (temp_end+temp_start)/2))
    
##RF ####
 #this is the best one   
      mod_rf_huds <- mgcv::gam(rate ~ s(temp, k = 5, by=curve_id), 
                          data = mutate(dplyr::filter(Sasaki2024_comb_d, metric == "RF", species == "hudsonica"), curve_id = as.factor(curve_id)), method="REML")

    lextr_rf_huds_all <- list()
    for(i in 1:length(unique(mod_rf_huds$model$curve_id))){
        lextr_hf_huds_i <- lextremagam::quantify_lextrema_multivar(mod = mod_rf_huds,
                                                                    var = "temp", 
                                                                    smooth = paste0("s(temp):curve_id", unique(mod_epr_huds$model$curve_id)[i]), 
                                                                    conf_level = 0.95, deriv_method = "gratia", frequentist = TRUE )
      lextr_rf_huds_all[[i]] <- lextr_hf_huds_i
      names(lextr_rf_huds_all)[i] <- as.character(unique(mod_hf_huds$model$curve_id)[i])
    }
    
    #extracting peaks 
    rf_huds_max <- purrr::map(lextr_rf_huds_all, ~ dplyr::filter(.x$segment_summary, feature %in% c("boundary_max", "local_max")))
    rf_huds_max <- purrr::map(rf_huds_max, ~ dplyr::mutate(.x, midpoint = (temp_end+temp_start)/2))



```

# DFO Shrimp Data #### 
Data from Interactive effects of predation and climate on the distributions of marine shellfish in the Northwest Atlantic
Sana Zabihi-Seissan, Krista D. Baker, Ryan R. E. Stanley, Tyler D. Tunney, Brittany Beauchamp, Hugues P. Benoit, David Brickman, Denis Chabot, Adam Cook, David Deslauriers, Mariano Koen-Alonso, Jake Lawlor, Arnault Le Bris, Darrell R. J. Mullowney, Marie-Julie Roux, Katherine R. Skanes, Zeliang Wang, Eric J. Pedersen
17 May 2024
```{r}
shrimp_dat <- read.csv("data-raw/real_data/diet_workgroup_trawl_data.csv")

#filtering for pandalus borealis (shrimp of fisheries interest) and NAFO divisions 2j, 3K, 3L (core areas of NFL shelf)
shrimp_dat <- filter(shrimp_dat, NAFO %in% c("3L", "2J", "3K"), year_surv>=2010, year_surv<=2015) %>%
  select(-Chionoecetes.opilio, -Gadus.morhua, -Hippoglossus.hippoglossus, -Reinhardtius.hippoglossoides, -Sebastes.mentella)
shrimp_dat$year_surv <- as.factor(shrimp_dat$year_surv)
shrimp_dat$NAFO <- as.factor(shrimp_dat$NAFO)
shrimp_dat$season <- as.factor(shrimp_dat$season)

shrimp_gam <- gam(Pandalus.borealis ~s(depth)+s(year_surv, bs="re")+s(NAFO, bs="re")+s(lat_dec), data=shrimp_dat, method="REML")

draw(shrimp_gam, residuals = FALSE)

shrimp_lextr <- lextremagam::quantify_lextrema_multivar(mod = shrimp_gam, var = "depth", 
                                                        smooth = "s(depth)", conf_level = 0.95, 
                                                        deriv_method = "gratia", frequentist = FALSE, step_size = 0.020)

plot_lextrema(shrimp_lextr)
lextremagam:::.plot_lextrema_multivar(shrimp_lextr)

shrimp_lextr$segment_summary
```

# Iarcella et al Feeding by temperature
Data from Figure 3 of Ecological impacts of invasive alien species along temperature gradients: testing the role of environmental matching
Josephine C. Iacarella, Jaimie T. A. Dick, Mhairi E. Alexander, Anthony Ricciardi
April 2015
```{r}
ias_data <- readr::read_xlsx(path = "Appendix B - Iacarella et al.xlsx", skip = 1)
ias_data <- ias_data[, (6):ncol(ias_data)]
colnames(ias_data) <- c("species", "temp_c", "max_feed_rate_pph")
ias_data$species <- as.factor(ias_data$species)
head(ias_data)

plot(ias_data$max_feed_rate_pph~ias_data$temp_c)

gam_ias_anomala <-gam(max_feed_rate_pph~s(temp_c, k=6),
              data = filter(ias_data, species == "Hemimysis anomala"), method = "REML", family=Gamma(link = "log"))
draw(gam_ias_anomala, residuals=TRUE)

gam_ias_diluviana<-gam(max_feed_rate_pph~s(temp_c, k=3),
                      data = filter(ias_data, species == "Mysis diluviana"), method = "REML", family=Gamma(link = "log"))
draw(gam_ias_diluviana, residuals=TRUE)

```
