---
title: "S05-create_plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{S05-create_plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lextremagam)
library(ggplot2)
library(dplyr)
```

#Import all sim data 
```{r}
setwd("~/Documents/Pedersen2023/Peaks/lextremagam")
rds_sim <- list.files("data-raw/sim/analy_results/", full.names = TRUE)
for (i in rds_sim ){
    readRDS(i)}

#merging earlier sim tests into a single summary list
e1_4_summary_ad <- list(e1 = e1_summary, 
                        e2 = e2_summary, 
                        e3 = e3_summary, 
                        e4 = e4_summary)
saveRDS(e1_4_summary_ad, "data-raw/sim/analy_results/e1_4_summary_adjalpha.RDS")
```

#Type I error
```{r}
plot_fpe <- function(summary, title){
  #TYPE I ERROR ####
  e1_fpe <- ggplot(data=tidyr::pivot_longer(summary$e1$fperror, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "fperror"))+
    geom_point(aes(x=dup, y=fperror, colour=method))+
    geom_line(aes(x=dup, y=fperror, colour=method))+
    geom_hline(aes(yintercept=0.05), linetype=2)+
    labs(title = paste0("Exp 1 FPE,  ", title))
  
e2_fpe <- ggplot(data=tidyr::pivot_longer(summary$e2$fperror, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "fperror"))+
  geom_point(aes(x=error, y=fperror, colour=method))+
  geom_line(aes(x=error, y=fperror, colour=method))+
  geom_hline(aes(yintercept=0.05), linetype=2)+
  labs(title = paste0("Exp 2 FPE,  ", title))

e4_fpe <- ggplot(data=tidyr::pivot_longer(summary$e4$fperror, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "fperror"))+
  geom_point(aes(x=param, y=fperror, colour=method))+
  geom_line(aes(x=param, y=fperror, colour=method))+
  geom_hline(aes(yintercept=0.05), linetype=2)+
  facet_grid(dup~func)+
  labs(title = paste0("Exp 4 FPE,  ", title))

return(list(e1_fpe = e1_fpe, e2_fpe = e2_fpe, e4_fpe = e4_fpe))
}

fpe_adaptive <- plot_fpe(e1_4_summary_ad, "Adaptive Spline, Alpha = 0.05")
fpe_tprs <- plot_fpe(e1_4_summary_tprs, "TPRS, Alpha = 0.05")
fpe_adaptive_adj <- plot_fpe(e1_4_summary_adjalpha_ad, "Adaptive Spline, Alpha = sqrt(0.05)")
fpe_tprs_adj <- plot_fpe(e1_4_summary_adjalpha, "TPRS, Alpha = sqrt(0.05)")


pdf("typeIerror_plots.pdf")
fpe_adaptive
fpe_tprs
fpe_adaptive_adj
fpe_tprs_adj
dev.off()

```

#Power
```{r}
plot_power <- function(summary, title){
    e1_p <- ggplot(data=tidyr::pivot_longer(summary$e1$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"))+
  geom_point(aes(x=dup, y=power, colour=method))+
  geom_line(aes(x=dup, y=power, colour=method))+
  labs(title = paste0("Exp 1 P,  ", title))
  
e2_p <- ggplot(data=tidyr::pivot_longer(summary$e2$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"))+
  geom_point(aes(x=error, y=power, colour=method))+
  geom_line(aes(x=error, y=power, colour=method))+
  labs(title = paste0("Exp 2 P,  ", title))

e3_p_1 <- ggplot(data=filter(tidyr::pivot_longer(summary$e3$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"), func!="var-p-width"))+
  geom_point(aes(x=peak, y=power, colour=method))+
  geom_line(aes(x=peak, y=power, colour=method))+
  facet_grid(dup~func)+
  labs(title = paste0("Exp 3 P,  ", title))
  
e3_p_2 <- ggplot(data=filter(tidyr::pivot_longer(e1_4_summary_adjalpha_ad$e3$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"), func=="var-p-width"))+
  geom_point(aes(x=param, y=power, colour=method))+
  geom_line(aes(x=param, y=power, colour=method))+
  facet_grid(dup~func)+
  labs(title = paste0("Exp 3 P,  ", title))

return(list(e1_fpe = e1_p, e2_p = e2_p, e3_p_1 = e3_p_1, e3_p_2 = e3_p_2))
}

p_adaptive <- plot_power(e1_4_summary_ad, "Adaptive Spline, Alpha = 0.05")
p_tprs <- plot_power(e1_4_summary_tprs, "TPRS, Alpha = 0.05")
p_adaptive_adj <- plot_power(e1_4_summary_adjalpha_ad, "Adaptive Spline, Alpha = sqrt(0.05)")
p_tprs_adj <- plot_power(e1_4_summary_adjalpha, "TPRS, Alpha = sqrt(0.05)")


pdf("power_plots.pdf")
p_adaptive
p_tprs
p_adaptive_adj
p_tprs_adj
dev.off()

```




pdf("e1_4_tprs_adj_ad_plots.pdf")
#POWER ####
ggplot(data=tidyr::pivot_longer(e1_4_summary_adjalpha_ad$e1$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"))+
  geom_point(aes(x=dup, y=power, colour=method))+
  geom_line(aes(x=dup, y=power, colour=method))

ggplot(data=tidyr::pivot_longer(e1_4_summary_adjalpha_ad$e2$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"))+
  geom_point(aes(x=error, y=power, colour=method))+
  geom_line(aes(x=error, y=power, colour=method))

ggplot(data=filter(tidyr::pivot_longer(e1_4_summary_adjalpha_ad$e3$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"), func!="var-p-width"))+
  geom_point(aes(x=peak, y=power, colour=method))+
  geom_line(aes(x=peak, y=power, colour=method))+
  facet_grid(dup~func)
ggplot(data=filter(tidyr::pivot_longer(e1_4_summary_adjalpha_ad$e3$power, cols = c("quad", "rprop", "pr_gr"), names_to = "method", values_to = "power"), func=="var-p-width"))+
  geom_point(aes(x=param, y=power, colour=method))+
  geom_line(aes(x=param, y=power, colour=method))+
  facet_grid(dup~func)


# COVERAGE ####
ggplot(data=e1_4_summary_adjalpha_ad$e1$coverage)+
  geom_point(aes(x=dup, y=coverage_gr))+
  geom_line(aes(x=dup, y=coverage_gr))+
  geom_hline(aes(yintercept=0.95), linetype=2)

ggplot(data=e1_4_summary_adjalpha_ad$e2$coverage)+
  geom_point(aes(x=error, y=coverage_gr))+
  geom_line(aes(x=error, y=coverage_gr))+
  geom_hline(aes(yintercept=0.95), linetype=2)

ggplot(data=filter(e1_4_summary_adjalpha_ad$e3$coverage, func!="var-p-width"))+
  geom_point(aes(x=peak, y=coverage_gr, colour=as.factor(dup)))+
  geom_line(aes(x=peak, y=coverage_gr, colour=as.factor(dup)))+
  geom_hline(aes(yintercept=0.95), linetype=2)+
  facet_wrap(~func)

ggplot(data=filter(e1_4_summary_adjalpha_ad$e3$coverage, func=="var-p-width"))+
  geom_point(aes(x=param, y=coverage_gr, colour=as.factor(dup)))+
  geom_line(aes(x=param, y=coverage_gr, colour=as.factor(dup)))+
  geom_hline(aes(yintercept=0.95), linetype=2)+
  facet_wrap(~func)

#PRECISION ####
ggplot(data=e1_4_summary_adjalpha_ad$e1$precision)+
  geom_point(aes(x=dup, y=width_avg_gr))+
  geom_line(aes(x=dup, y=width_avg_gr))+
  geom_ribbon(aes(x = dup, ymin=width_avg_gr-width_sd_gr, ymax=width_avg_gr+width_sd_gr), alpha=0.2)

ggplot(data=e1_4_summary_adjalpha_ad$e2$precision)+
  geom_point(aes(x=error, y=width_avg_gr))+
  geom_line(aes(x=error, y=width_avg_gr))+
  geom_ribbon(aes(x = error, ymin=width_avg_gr-width_sd_gr, ymax=width_avg_gr+width_sd_gr), alpha=0.2)

ggplot(data=filter(e1_4_summary_adjalpha_ad$e3$precision, func !="var-p-width"))+
  geom_point(aes(x=peak, y=width_avg_gr, colour=as.factor(dup)))+
  geom_line(aes(x=peak, y=width_avg_gr, colour=as.factor(dup)))+
  geom_ribbon(aes(x = peak, ymin=width_avg_gr-width_sd_gr, ymax=width_avg_gr+width_sd_gr, fill=as.factor(dup)), alpha=0)+
  facet_wrap(~func)

ggplot(data=filter(e1_4_summary_adjalpha_ad$e3$precision, func =="var-p-width"))+
  geom_point(aes(x=param, y=width_avg_gr, colour=as.factor(dup)))+
  geom_line(aes(x=param, y=width_avg_gr, colour=as.factor(dup)))+
  geom_ribbon(aes(x = param, ymin=width_avg_gr-width_sd_gr, ymax=width_avg_gr+width_sd_gr, fill=as.factor(dup)), alpha=0)+
  facet_wrap(~func)

# BIAS ####
ggplot(data=filter(e1_4_summary_adjalpha_ad$e1$bias, func== "norm"))+
  geom_point(aes(x=dup, y=bias_avg_gr))+
  geom_line(aes(x=dup, y=bias_avg_gr))+
  geom_ribbon(aes(x = dup, ymin=bias_avg_gr-bias_sd_gr, ymax=bias_avg_gr+bias_sd_gr), alpha=0.2)+
  geom_hline(aes( yintercept=0), linetype=2)

ggplot(data=filter(e1_4_summary_adjalpha_ad$e2$bias, func== "norm"))+
  geom_point(aes(x=error, y=bias_avg_gr))+
  geom_line(aes(x=error, y=bias_avg_gr))+
  geom_ribbon(aes(x = error, ymin=bias_avg_gr-bias_sd_gr, ymax=bias_avg_gr+bias_sd_gr), alpha=0.2)+
  geom_hline(aes(yintercept=0), linetype=2)

ggplot(data=filter(e1_4_summary_adjalpha_ad$e3$bias, func!= "var-p-width"))+
  geom_point(aes(x=peak, y=bias_avg_gr, colour=as.factor(dup)))+
  geom_line(aes(x=peak, y=bias_avg_gr, colour=as.factor(dup)))+
  geom_ribbon(aes(x = peak, ymin=bias_avg_gr-bias_sd_gr, ymax=bias_avg_gr+bias_sd_gr, fill=as.factor(dup)), alpha=0.1)+
  geom_hline(aes( yintercept=0), linetype=2)+
facet_wrap(~func)

ggplot(data=filter(e1_4_summary_adjalpha_ad$e3$bias, func== "var-p-width"))+
  geom_point(aes(x=param, y=bias_avg_gr, colour=as.factor(dup)))+
  geom_line(aes(x=param, y=bias_avg_gr, colour=as.factor(dup)))+
  geom_ribbon(aes(x = param, ymin=bias_avg_gr-bias_sd_gr, ymax=bias_avg_gr+bias_sd_gr, fill=as.factor(dup)), alpha=0.1)+
  geom_hline(aes( yintercept=0), linetype=2)+
facet_wrap(~func)

dev.off()



